\chapter{Background}
\label{chap:background}

\section{Physically Based Rendering}

\subsection{The Rendering Equation}

The rendering equation, introduced by Kajiya~\cite{Kajiya1986}, describes
the equilibrium distribution of light in a scene:
\begin{equation}
  L_o(\mathbf{x}, \omega_o) =
    L_e(\mathbf{x}, \omega_o) +
    \int_{\Omega}
      f_r(\mathbf{x}, \omega_i, \omega_o)\,
      L_i(\mathbf{x}, \omega_i)\,
      (\omega_i \cdot \mathbf{n})\,
    \mathrm{d}\omega_i
  \label{eq:rendering}
\end{equation}
where $L_o$ is outgoing radiance, $L_e$ is emitted radiance,
$f_r$ is the BRDF, and $\Omega$ is the hemisphere of incoming directions.

\subsection{Monte Carlo Integration}

\todo{Summarize MC integration, importance sampling, MIS (Veach 1997).}

\subsection{Path Tracing}

\todo{Algorithm overview; Kajiya 1986, Veach 1997.}

\section{GPU Architecture}

\subsection{Tile-Based Deferred Rendering (TBDR)}

\todo{Explain TBDR: PowerVR, Apple GPU, ARM Mali, Vivante.
Morton encoding. Bandwidth reduction.}

\subsection{Ray Tracing Hardware}

\todo{BVH, traversal units, NVIDIA RT Core, Intel Xe,
      RPU (Woop et al.\ 2005).}

\section{Open-Source Hardware}

\subsection{RISC-V Ecosystem}

\todo{Brief overview of open ISA movement.}

\subsection{Chisel and High-Level HDL}

\todo{Chisel/FIRRTL, generators, type-safe HDL.}

\subsection{Related Open-Source GPUs}

\begin{description}
  \item[Nyuzi] GPGPU vector processor by Jeff Bush. No tapeout.
  \item[Vortex] RISC-V based GPGPU from Georgia Tech. Research only.
  \item[\borg{}] (this work) TBDR GPU with physical tapeout.
\end{description}
